\LARGE \bf Geometric Adaptive Control of Attitude Dynamics on $\SO$\\ with State Inequality Constraints

### 2015-09-23 ACC16.2.tex
starting on simulation section

\begin{proof}
	See~\Cref{proof:config_error}
\end{proof}

@@

\begin{proof}
	See~\Cref{proof:error_dyn}
\end{proof}

@@

\begin{proof}\label{proof:att_control}
	Consider the following Lyapunov function:
	\begin{gather}
		V = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R \Psi(R,R_d) \, . \label{eqn:v_nodist}
	\end{gather}
	From~\cref{item:prop_psi_psd}, we obtain
	\begin{gather*}
		V \geq 0 \, .
	\end{gather*}
	Using~\cref{eqn:eW_dot,eqn:psi_dot} with \( \Delta = 0 \), the time derivative of \( V \) is given by
	\begin{align}
		\dot{V} &= e_\Omega^T \parenth{ u - \Omega \times J \Omega } + k_R e_R \cdot e_\Omega \nonumber \\
		&= -k_\Omega \norm{e_\Omega}^2 \, . \label{eqn:vdot_nodist}
	\end{align}
	Since \( V \) is positive definite and \( \dot{V} \) is negative definite this implies that the zero equilibrium point \( e_R, e_\Omega \) is asymptotically stable.
\end{proof}

@@

\begin{prop}[Adaptive Attitude Control]\label{prop:adaptive_control}
Given  a desired attitude command \( R_d, \Omega_d = 0 \) and positive constants \( k_R, k_\Omega, k_\Delta, c \in \R \), we define a control input \( u \in \R^3\) and an adaptive update law for \( \bar{\Delta} \) as follows:
\begin{align}
	u &= - k_R e_R - k_\Omega e_\Omega + \Omega \times J \Omega - W \bar{\Delta} \, , \label{eqn:adaptive_control} \\
	\dot{\bar{\Delta}} &= k_\Delta W^T \parenth{e_\Omega + c e_R} \, . \label{eqn:delta_dot}
\end{align}
If \( c \) is chosen such that
\begin{gather}
	0 < c < \frac{4 k_R k_\Omega}{k_\Omega^2 + 4 k-R \lambda_M H} \, , \label{eqn:c_bound}
\end{gather}
  the zero equilibrium of the error vectors \( e_R, e_\Omega\) are asymptotically stable.
\end{prop}
\begin{proof}[Proof of~\cref{prop:adaptive_control}]\label{proof:adaptive_control}
	Consider the Lyapunov function \( V \) given by
	\begin{gather}
		V = \frac{1}{2} e_\Omega J e_\Omega + k_R \Psi + c J e_\Omega^T e_R + \frac{1}{2 k_\Delta} e_\Delta^T e_\Delta \, , \label{eqn:v_adapt}
	\end{gather}
	over the domain \( D \) in~\cref{eqn:domain}.
	From~\cref{prop:eR_dot_bound}, the Lyapunov function is bounded in \( D \) by
	\begin{gather}
		V \leq z^T W z \, , \label{eqn:v_upper_bound}
	\end{gather}
	where \( e_\Delta = \Delta - \bar{\Delta} \), \( z = \begin{bmatrix}{ccc}e_R & e_\Omega & e_\Delta\end{bmatrix} \) and the matrix \(W \in \R^{3 \times 3}\) is given by
	\begin{gather*}
		W = \begin{bmatrix}{ccc}
			k_R \psi & \frac{1}{2} c \lambda_M & 0 \\
			\frac{1}{2} c \lambda_M & \frac{1}{2} \lambda_M & 0 \\
			0 & 0 & \frac{1}{2 k_\Delta}
		\end{bmatrix} \, .
	\end{gather*}
	The time derivative of with the control inputs~\cref{eqn:adaptive_control} is defined as
	\begin{align}
		\dot{V} =& - k_\Omega e_\Omega^T e_\Omega + \parenth{e_\Omega + c e_R}^T W e_\Delta - k_R c e_R^T e_R \nonumber\\
		&- k_\Omega c e_R^T e_\Omega + c J e_\Omega^T \dot{e}_R - \frac{1}{k_\Delta} e_\Delta^T \dot{\bar{\Delta}} \, , \label{eqn:vdot}
	\end{align}
	where we used \( \dot{e}_\Delta = - \dot{\bar{\Delta}} \).
	The terms linearly dependent on \( e_\Delta\) are combined with~\cref{eqn:delta_dot} to yield
	\begin{align*}
		 e_\Delta^T \parenth{W^T \parenth{e_\Omega + c e_R} - \frac{1}{k_\Delta} \dot{\bar{\Delta}}} = 0 \, . 
	\end{align*}
	Using~\Cref{prop:eR_dot_bound} an upper bound on \( \dot{V} \) is written as
	\begin{gather*}
		\dot{V} \leq -z^T M z \, ,
	\end{gather*}
	where the matrix \( M \in \R^{2 \times 2} \) is 
	\begin{gather}
		M = \begin{bmatrix}{cc}
			k_R c & \frac{k_\Omega c}{2} \\
			\frac{k_\Omega c}{2} & k_\Omega - c \lambda_M H
		\end{bmatrix} \, .
	\end{gather}
	If \( c \) is chosen such that~\cref{eqn:c_bound} is satisfied the matrix \( M \) is positive definite.
	This implies that the zero equilibrium \( e_R, e_\Omega\) is asymptotically stable. 
\end{proof}

Local asymptotic stability

Domain avoids undesired equilibria

Mention how for a fixed disturbance this is integral control

Go through the various developments of proving the disturbance case is stable for correct choice of gains

Positive semi definite matrix

Give example for a simple case

\section{Numerical Examples}
Parameters used in the simulation

Inertia matrix of hexrotor

control gains

define sensor and constraint axis

Define initial and desired condition

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\subsection{Proof of~\Cref{prop:config_error}} \label{proof:config_error}
The proof of~\cref{item:prop_era_upbound} is available in~\cite{lee2011a}.
To prove~\cref{item:prop_psi_psd} we note that~\cref{eqn:A} is a locally positive definite function about \( R = R_d \)~\cite{bullo2004}.
The constraint angle is assumed \( \ang{0} \leq \theta \leq \ang{90} \) such that \( 0 \leq \cos \theta \).
The term \( r^T R^T v \) represents the cosine of the angle between the body fixed vector \( r \) and the inertial vector \( v \). 
It follows that
\begin{align*}
	0 \leq  \frac{\cos \theta -  r^T R^T v}{1 + \cos \theta} \leq 1 \, ,
\end{align*}
for all \( R \in SO \). 
As a result, the negative logarithm is always positive and
\begin{align*}
	1 < B \, .
\end{align*}
The product~\cref{eqn:psi} is composed of two positive terms and is therefore also positive definite, and has a single critical point at \( R = R_d \).

To show~\cref{item:prop_crit}, we determine the critical points of~\cref{eqn:eRA}.
The attitude error between \( R \) and \( R_d \) is defined as \( Q \in SO \) as
\begin{align*}
	Q = R_d^T R = 
	\begin{bmatrix}
		q_{11} & q_{12} & q_{13} \\
		q_{21} & q_{22} & q_{23} \\
		q_{31} & q_{32} & q_{33}
	\end{bmatrix} \, .
\end{align*}
Equating \( e_{R_{A}} = 0 \) yields the following equivalent relations
\begin{subequations}
\begin{align}
	g_3 q_{32} &= g_2 q_{23} \, , \\
	g_1 q_{13} &= g_3 q_{31} \, , \\
	g_2 q_{21} &= g_1 q_{12} \, .
\end{align}
\end{subequations}
These relations allow one to redefine the attitude error \( Q \) in terms of the independent upper diagonal elements.
Noting that \( Q \in \SO \) is an orthonormal matrix, with unit length row and columns gives
\begin{subequations}
\begin{align}
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 + \parenth{1 - \frac{g_1^2}{g_3^2}}q_{13}^2 &= 0 \, , \\
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 - \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, , \\
	\parenth{1-\frac{g_1^2}{g_3^2}} q_{13}^2 + \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, .
\end{align}
\end{subequations}
As a result, all the off diagonal terms of \( Q = R_d^T R \), namely \( q_{12}= q_{13} = q_{23} = 0 \).
Since \( Q \in \SO \), the diagonal elements \( q_{11}, q_{22}, \text{ and } q_{33} \) are either \( +1 \) or \(-1\).
There are four critical points of \( e_{R_A} \), the desired attitude \( R_d \) as well as rotations about each body fixed axis by \ang{180}.
The repulsive error vector \( e_{R_{B}} \) is zero only when the numerator \( \parenth{R^T v}^\wedge r = 0 \). 
This condition only occurs if the desired attitude results in the body fixed vector \( r \) becoming aligned with \(R^T v \) while simultaneously satisfying \(\braces{R_d} \cup \braces{R_d \exp(\pi \hat{s}} \) for \( s \in \braces{e_1, e_2, e_3} \) .
The desired equilibrium is \( e_R = 0 \) and \( A = 0\).
\subsection{Proof of~\Cref{prop:error_dyn}}\label{proof:error_dyn}
From the kinematics~\cref{eqn:Rdot} and noting that \( \dot{R}_d = 0 \) the time derivative of \( R_d^T R \) is given as
\begin{gather*}
	\diff{}{t} \parenth{R_d^T R} = R_d^T R \hat{e}_\Omega \, .
\end{gather*}
Applying this to the time derivative of~\cref{eqn:A} gives
\begin{gather*}
	\diff{}{t} A = -\frac{1}{2} \tr{G R_d^T R \hat{e}_\Omega} \, .
\end{gather*}
Applying~\cref{eqn:hat1} into this shows~\cref{eqn:A_dot}.
Next, the time derivative of the repulsive error function is given by
\begin{gather*}
	\diff{}{t} B = \frac{r_i^T \parenth{\hat{\Omega} R^T} v_i}{\alpha \parenth{r_i^T R^T v_i - \cos \theta}} \, .
\end{gather*}
Using the scalar triple product, given by~\cref{eqn:STP}, one can reduce this to~\cref{eqn:B_dot}.
The time derivative of the attractive attitude error vector, \( e_{R_A} \), is given by
\begin{gather*}
	\diff{}{t} e_{R_A} = \frac{1}{2} \parenth{\hat{e}_\Omega R^T R_d G + (R^T R_d G)^T \hat{e}_\Omega}^\vee \, .
\end{gather*}
Using the hat map property given in~\cref{eqn:xAAx} this is further reduced to~\cref{eqn:eRA_dot,eqn:E}.

The time-derivative of the angular velocity error, \( e_\Omega \), is given by
\begin{align*}
	\diff{}{t} e_\Omega &= \dot{\Omega} \\
	&= J^{-1} \parenth{u - \Omega \times J \Omega + W(R,\Omega) \Delta } \, ,
\end{align*}
where we have used the dynamic equations of motion from~\cref{eqn:Wdot}. 
\subsection{Proof of }
Stability proofs for case with and without disturbance

>>>>

\begin{proof}[Proof of~\Cref{prop:config_error}]\label{proof:config_error}
The proof of~\cref{item:prop_era_upbound} is available in~\cite{lee2011a}.
To prove~\cref{item:prop_psi_psd} we note that~\cref{eqn:A} is a locally positive definite function about \( R = R_d \)~\cite{bullo2004}.
The constraint angle is assumed \( \ang{0} \leq \theta \leq \ang{90} \) such that \( 0 \leq \cos \theta \).
The term \( r^T R^T v \) represents the cosine of the angle between the body fixed vector \( r \) and the inertial vector \( v \). 
It follows that
\begin{align*}
	0 \leq  \frac{\cos \theta -  r^T R^T v}{1 + \cos \theta} \leq 1 \, ,
\end{align*}
for all \( R \in SO \). 
As a result, the negative logarithm is always positive and
\begin{align*}
	1 < B \, .
\end{align*}
The product~\cref{eqn:psi} is composed of two positive terms and is therefore also positive definite, and has a single critical point at \( R = R_d \).

To show~\cref{item:prop_crit}, we determine the critical points of~\cref{eqn:eRA}.
The attitude error between \( R \) and \( R_d \) is defined as \( Q \in SO \) as
\begin{align*}
	Q = R_d^T R = 
	\begin{bmatrix}
		q_{11} & q_{12} & q_{13} \\
		q_{21} & q_{22} & q_{23} \\
		q_{31} & q_{32} & q_{33}
	\end{bmatrix} \, .
\end{align*}
Equating \( e_{R_{A}} = 0 \) yields the following equivalent relations
\begin{subequations}
\begin{align}
	g_3 q_{32} &= g_2 q_{23} \, , \\
	g_1 q_{13} &= g_3 q_{31} \, , \\
	g_2 q_{21} &= g_1 q_{12} \, .
\end{align}
\end{subequations}
These relations allow one to redefine the attitude error \( Q \) in terms of the independent upper diagonal elements.
Noting that \( Q \in \SO \) is an orthonormal matrix, with unit length row and columns gives
\begin{subequations}
\begin{align}
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 + \parenth{1 - \frac{g_1^2}{g_3^2}}q_{13}^2 &= 0 \, , \\
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 - \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, , \\
	\parenth{1-\frac{g_1^2}{g_3^2}} q_{13}^2 + \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, .
\end{align}
\end{subequations}
As a result, all the off diagonal terms of \( Q = R_d^T R \), namely \( q_{12}= q_{13} = q_{23} = 0 \).
Since \( Q \in \SO \), the diagonal elements \( q_{11}, q_{22}, \text{ and } q_{33} \) are either \( +1 \) or \(-1\).
There are four critical points of \( e_{R_A} \), the desired attitude \( R_d \) as well as rotations about each body fixed axis by \ang{180}.
The repulsive error vector \( e_{R_{B}} \) is zero only when the numerator \( \parenth{R^T v}^\wedge r = 0 \). 
This condition only occurs if the desired attitude results in the body fixed vector \( r \) becoming aligned with \(R^T v \) while simultaneously satisfying \(\braces{R_d} \cup \braces{R_d \exp(\pi \hat{s}} \) for \( s \in \braces{e_1, e_2, e_3} \) .
The desired equilibrium is \( e_R = 0 \) and \( A = 0\).
\end{proof}

@@

\begin{proof}[Proof of~\Cref{prop:error_dyn}]\label{proof:error_dyn}
From the kinematics~\cref{eqn:Rdot} and noting that \( \dot{R}_d = 0 \) the time derivative of \( R_d^T R \) is given as
\begin{gather*}
	\diff{}{t} \parenth{R_d^T R} = R_d^T R \hat{e}_\Omega \, .
\end{gather*}
Applying this to the time derivative of~\cref{eqn:A} gives
\begin{gather*}
	\diff{}{t} A = -\frac{1}{2} \tr{G R_d^T R \hat{e}_\Omega} \, .
\end{gather*}
Applying~\cref{eqn:hat1} into this shows~\cref{eqn:A_dot}.
Next, the time derivative of the repulsive error function is given by
\begin{gather*}
	\diff{}{t} B = \frac{r_i^T \parenth{\hat{\Omega} R^T} v_i}{\alpha \parenth{r_i^T R^T v_i - \cos \theta}} \, .
\end{gather*}
Using the scalar triple product, given by~\cref{eqn:STP}, one can reduce this to~\cref{eqn:B_dot}.
The time derivative of the attractive attitude error vector, \( e_{R_A} \), is given by
\begin{gather*}
	\diff{}{t} e_{R_A} = \frac{1}{2} \parenth{\hat{e}_\Omega R^T R_d G + (R^T R_d G)^T \hat{e}_\Omega}^\vee \, .
\end{gather*}
Using the hat map property given in~\cref{eqn:xAAx} this is further reduced to~\cref{eqn:eRA_dot,eqn:E}.

The time-derivative of the angular velocity error, \( e_\Omega \), is given by
\begin{align*}
	\diff{}{t} e_\Omega &= \dot{\Omega} \\
	&= J^{-1} \parenth{u - \Omega \times J \Omega + W(R,\Omega) \Delta } \, ,
\end{align*}
where we have used the dynamic equations of motion from~\cref{eqn:Wdot}. 
\end{proof}

@@

\begin{proof}[Proof of~\Cref{prop:att_control}]\label{proof:att_control}
	Consider the following Lyapunov function:
	\begin{gather}
		V = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R \Psi(R,R_d) \, . \label{eqn:v_nodist}
	\end{gather}
	From~\cref{item:prop_psi_psd}, we obtain
	\begin{gather*}
		V \geq 0 \, .
	\end{gather*}
	Using~\cref{eqn:eW_dot,eqn:psi_dot} with \( \Delta = 0 \), the time derivative of \( V \) is given by
	\begin{align}
		\dot{V} &= e_\Omega^T \parenth{ u - \Omega \times J \Omega } + k_R e_R \cdot e_\Omega \nonumber \\
		&= -k_\Omega \norm{e_\Omega}^2 \, . \label{eqn:vdot_nodist}
	\end{align}
	Since \( V \) is positive definite and \( \dot{V} \) is negative definite this implies that the zero equilibrium point \( e_R, e_\Omega \) is asymptotically stable.
\end{proof}

@@

\begin{prop}[Adaptive Attitude Control]\label{prop:adaptive_control}
Given  a desired attitude command \( R_d, \Omega_d = 0 \) and positive constants \( k_R, k_\Omega, k_\Delta, c \in \R \), we define a control input \( u \in \R^3\) and an adaptive update law for \( \bar{\Delta} \) as follows:
\begin{align}
	u &= - k_R e_R - k_\Omega e_\Omega + \Omega \times J \Omega - W \bar{\Delta} \, , \label{eqn:adaptive_control} \\
	\dot{\bar{\Delta}} &= k_\Delta W^T \parenth{e_\Omega + c e_R} \, . \label{eqn:delta_dot}
\end{align}
If \( c \) is chosen such that
\begin{gather}
	0 < c < \frac{4 k_R k_\Omega}{k_\Omega^2 + 4 k-R \lambda_M H} \, , \label{eqn:c_bound}
\end{gather}
  the zero equilibrium of the error vectors \( e_R, e_\Omega\) are asymptotically stable.
\end{prop}
\begin{proof}[Proof of~\cref{prop:adaptive_control}]\label{proof:adaptive_control}
	Consider the Lyapunov function \( V \) given by
	\begin{gather}
		V = \frac{1}{2} e_\Omega J e_\Omega + k_R \Psi + c J e_\Omega^T e_R + \frac{1}{2 k_\Delta} e_\Delta^T e_\Delta \, , \label{eqn:v_adapt}
	\end{gather}
	over the domain \( D \) in~\cref{eqn:domain}.
	From~\cref{prop:eR_dot_bound}, the Lyapunov function is bounded in \( D \) by
	\begin{gather}
		V \leq z^T W z \, , \label{eqn:v_upper_bound}
	\end{gather}
	where \( e_\Delta = \Delta - \bar{\Delta} \), \( z = \begin{bmatrix}e_R & e_\Omega & e_\Delta\end{bmatrix} \) and the matrix \(W \in \R^{3 \times 3}\) is given by
	\begin{gather*}
		W = \begin{bmatrix}
			k_R \psi & \frac{1}{2} c \lambda_M & 0 \\
			\frac{1}{2} c \lambda_M & \frac{1}{2} \lambda_M & 0 \\
			0 & 0 & \frac{1}{2 k_\Delta}
		\end{bmatrix} \, .
	\end{gather*}
	The time derivative of with the control inputs~\cref{eqn:adaptive_control} is defined as
	\begin{align}
		\dot{V} =& - k_\Omega e_\Omega^T e_\Omega + \parenth{e_\Omega + c e_R}^T W e_\Delta - k_R c e_R^T e_R \nonumber\\
		&- k_\Omega c e_R^T e_\Omega + c J e_\Omega^T \dot{e}_R - \frac{1}{k_\Delta} e_\Delta^T \dot{\bar{\Delta}} \, , \label{eqn:vdot}
	\end{align}
	where we used \( \dot{e}_\Delta = - \dot{\bar{\Delta}} \).
	The terms linearly dependent on \( e_\Delta\) are combined with~\cref{eqn:delta_dot} to yield
	\begin{align*}
		 e_\Delta^T \parenth{W^T \parenth{e_\Omega + c e_R} - \frac{1}{k_\Delta} \dot{\bar{\Delta}}} = 0 \, . 
	\end{align*}
	Using~\Cref{prop:eR_dot_bound} an upper bound on \( \dot{V} \) is written as
	\begin{gather*}
		\dot{V} \leq -z^T M z \, ,
	\end{gather*}
	where the matrix \( M \in \R^{2 \times 2} \) is 
	\begin{gather}
		M = \begin{bmatrix}
			k_R c & \frac{k_\Omega c}{2} \\
			\frac{k_\Omega c}{2} & k_\Omega - c \lambda_M H
		\end{bmatrix} \, .
	\end{gather}
	If \( c \) is chosen such that~\cref{eqn:c_bound} is satisfied the matrix \( M \) is positive definite.
	This implies that the zero equilibrium \( e_R, e_\Omega\) is asymptotically stable. 
	The domain \( D \) does that include the undesired equilibria and ensures that as \( e_R \to 0 \) the attitude also converges \( R \to R_d \).
\end{proof}
Nonlinear adaptive controllers have been developed for attitude stabilization in terms of modified Rodriguez parameters and quaternions, as well as attitude tracking in terms of Euler angles. 
The proposed control system is developed on \(\SO\) and avoids the singularities of Euler angles and Rodriguez parameters while incorporating state inequality constraints. 
In addition, the unwinding and double coverage ambiguity of quaternions are also completely avoided.

\section{Numerical Examples}
We demonstrate the performance of the proposed control system via numerical simulation.
The inertia tensor of a rigid body is given as
\begin{gather*}
	J = \begin{bmatrix}
	\num{5.571e-3} & \num{6.177e-5} & \num{-2.503e-5} \\
	\num{6.177e-5} & \num{5.571e-3} & \num{1.001e-5} \\
	\num{-2.503e-5} & \num{1.007e-5} & \num{1.051e-2}
	\end{bmatrix} \, .
\end{gather*} 
The control system parameters are chosen as
\begin{gather*}
	G = \text{diag} \parenth{0.9,1.1,1.0}\, , \quad k_R = 0.4 \, , \quad	k_\Omega = 0.296 \, ,\\
	c = 1.0 \, , \quad k_\Delta = 0.5 \, , \quad \alpha = 10 \, .
\end{gather*}
A body fixed sensor is defined as
\begin{gather*}
	r = \begin{bmatrix} 1 & 0 & 0 \end{bmatrix} \, ,
\end{gather*}
while multiple inequality constraints are defined in~\cref{tab:constraints}.

The initial and desired states are defined using Euler angles as~\cite{shuster1993},
\begin{gather*}
	R(0) = I \, , \quad \Omega(0) = 0 \, ,\\
	R_d = ROT3(\ang{90}) \, , \quad \Omega_d = 0 \, .
\end{gather*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

\subsection{Proof of }
Stability proofs for case with and without disturbance
